---
title: "Analysis Immigrant Experiment Data with OLS, Logistic Model and Conditional Logistic Model, Plus Clustering"
author: "Wei-Lun, Lin"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)  # data processing
library(broom)
library(kableExtra)  # print table
library(survival)
library(cjoint)
library(marginaleffects)

library(numDeriv)
library(sandwich)
library(lmtest)
library(multiwayvcov)

source("./0-code/functions.R")
```

-   背景: 以 OLS, logistic regression, conditional logistic regression 分析 Hainmueller et al. (2014) 的 immigrant experiment 資料，並於此文件進一步比較 1. 有無加入交互作用、2. 有無對受試者層次做 cluster-robust standand error 對於估計上的差異。
<!-- -   結果: 比較 logistic 和 conditional logistic，發現 -->
<!-- ... OLS 得到的 AMCE 和 logistic regression 得到的估計非常接近，而 logistic 和 conditional logistic regression 的估計結果也相差不大。此外，分別比較 logistic 和 conditional logistic regression 可以發現，控制 cluster-robust standard error 前後得到的 standard error 也很類似，此結果和實驗的隨機生成 profiles 機制息息相關。 -->
<!-- -   未來工作:... 再加入無母數的方法，先對受試者的潛在分群特徵做分群，再用模型分別 fit 資料，並比較事先和事後控制 cluster 所得到的表現。 -->


## Immigrant Experiment

```{r import-data-immigrant}
immigrant <- read_rds("../1-data/2-processed/immigrant_data.rds")
immigrant <- immigrant %>% 
  mutate(
    id = row_number(),
    alt_id = id + 3 - 2 * profile
  )

immigrant %>% 
  head() %>% 
  kable("html", caption="First 6 Rows of immigrant Data") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>% 
  scroll_box(width = "100%")
```

在 immigrant experiment 中，immigrant 有9個 attributes:

| Attribute | Level 1 | Level 2 | Level 3 | Level 4 | Level 5 | Level 6 |  Level 7 |  Level 8 |  Level 9 |  Level 10 |  Level 11 |
|-----------|--------|---------------|--------|--------|-------------|-------------|--------|--------|--------|--------|---------|
| Education (7) | No formal education | Equivalent to completing fourth grade in the US | Equivalent to completing eighth grade in the US | Equivalent to completing high school in the US | Equivalent to completing two years of college in the US | Equivalent to completing a college degree in the US | Equivalent to completing a graduate degree in the US | \- | \- | \- | \- |
| Gender (2) | female | male | \- | \- | \- | \- | \- | \- | \- | \- | \- |
| Origin (10) | Germany | France | Mexico | Philippines | Poland | India | China | Sudan | Somalia | Iraq | \- |
| Application Reason (3) | Reunite with family members already in the U.S. | Seek better job in U.S. | Escape political/religious persecution | \- | \- | \- | \- | \- | \- | \- | \- |
| Profession (11) | Janitor | Waiter | Child care provider | Gardener | Financial analyst | Construction worker | Teacher | Computer programmer | Nurse | Research scientist | Doctor |
| Job experience (4) | No job training or prior experience | One or two years of job training and experience | Three to five years of job training and experience | More than five years of job training and experience | \- | \- | \- | \- | \- | \- | \- |
| Job plans (4) | Has a contract with a U.S. employer | Does not have a contract with a U.S. employer but has done job interviews | Will look for work after arriving in the U.S. | Has no plans to look for work at this time | \- | \- | \- | \- | \- | \- | \- |
| Prior trips to US (5) | Never been to the U.S. | Entered U.S. once before on a tourist visa | Has visited the U.S. many times before on tourist visas | Spent six months with family members in the U.S | Entered the U.S. once before without legal authorization | \- | \- | \- | \- | \- | \- |
| Language (4) | During the admission interview, this applicant spoke fluent English | During admission interview, this applicant spoke broken English | During admission interview, this applicant tried to speak English but was unable | During admission interview, this applicant spoke [language] and used an interpreter | \- | \- | \- | \- | \- | \- | \- |

排除不合理的 profiles 後，共有 $936960$ 種 potential profiles。

此外，為了方便後續圖表呈現，先對某些 attributes 的 levels 重新命名。

```{r level-rename}
immigrant <- immigrant %>% 
  mutate(
    Education. = as.factor(paste0("Education.", as.integer(Education.))),
    ApplicationReason. = as.factor(paste0("ApplicationReason.", as.integer(ApplicationReason.))),
    JobExperience. = as.factor(paste0("JobExperience.", as.integer(JobExperience.))),
    JobPlans. = as.factor(paste0("JobPlans.", as.integer(JobPlans.))),
    PriorTripstoUS. = as.factor(paste0("PriorTripstoUS.", as.integer(PriorTripstoUS.))),
    Language. = as.factor(paste0("Language.", as.integer(Language.)))
  )
```



## Model Fitting

因為在此實驗設計中，attributes 之間並不完全獨立（completely independent randomization），
而是會因為現實限制而導致某些 profiles 不應該出現，例如：

-  一部份的工作 （`profession`）會和教育（`education`）有關
-  一部分移民的理由（`application reason`）會和原國家（`origin`）有關

在此種情形下，attributes 之間不再是完全獨立，而是 conditionally independent randomization。
即儘管某個 attribute A 和另一 attribute B 相關，但在給定該 attribute B 的情況下，
attribute A 的分配會獨立於其他所有 attributes。
因此不能直接用只有主要效果的 models，而是要在模型中加入交互作用。

```{r res-all}
est.all.immigrant <- data.frame()
```

```{r design}
immigrant.design <- makeDesign(
  type="constraints",
  attribute.levels = list(
    "Education." = sort(unique(immigrant$Education.)),
    "Gender." = sort(unique(immigrant$Gender.)),
    "Origin." = sort(unique(immigrant$Origin.)),
    "ApplicationReason." = sort(unique(immigrant$ApplicationReason.)),
    "Profession." = sort(unique(immigrant$Profession.)),
    "JobExperience." = sort(unique(immigrant$JobExperience.)),
    "JobPlans." = sort(unique(immigrant$JobPlans.)),
    "PriorTripstoUS." = sort(unique(immigrant$PriorTripstoUS.)),
    "Language." = sort(unique(immigrant$Language.))
  ),
  constraints = list(
    list(
      "Education." = c("Education.1", "Education.2", "Education.3", "Education.4"),
      "Profession." = c("Financial analyst", "Computer programmer", "Doctor", "Research scientist")
    ),
    list(
      "ApplicationReason." = c("ApplicationReason.3"),
      "Origin." = c("Germany", "France", "Mexico", "Philippines", "Poland", "India")
    )
  )
)

form.immigrant <- select ~ Education. + Gender. + Origin. + ApplicationReason. + Profession. + JobExperience. + JobPlans. + PriorTripstoUS. + Language.
form_int.immigrant <- select ~ Gender. + JobExperience. + JobPlans. + PriorTripstoUS. + Language. + Education. * Profession. + Origin. * ApplicationReason.
```



### immigrant: Linear Regression

根據 Hainmueller et al. (2014)，
可以用 ordinary least squares (OLS) regression 來估計 attributes 各個 levels 的 average marginal component effects (AMCE)，
即平均掉其他 attributes 的影響後，該 level 的邊際效果。
在 immigrant experiment 這樣的 2AFC 的 task 中，估計出來的數值為「該 term 出現時，選項被選擇機率的增減」。

```{r ols}
# OLS without interaction
ols.immigrant <- fit_model(form.immigrant, dat = immigrant, model = "ols")
est.all.immigrant <- ols.immigrant[[2]] %>% 
  select(-c(statistic, p.value)) %>% 
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error
  ) %>% 
  rename_with(~ paste0(.x, ".OLS"), .cols = -c(attribute, term))

# OLS with clustering and without interaction
ols_clus.immigrant <- fit_model(form.immigrant, dat = immigrant, model = "ols", cluster = "subject")
est.all.immigrant <- ols_clus.immigrant[[2]] %>% 
  select(-c(statistic, p.value)) %>% 
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error
  ) %>% 
  rename_with(~ paste0(.x, ".OLS_clustered"), .cols = -c(attribute, term)) %>% 
  left_join(est.all.immigrant, ., by = join_by(attribute, term))

# OLS with interaction
# this part is completed with the `amce` function in the `cjoint` package
ols_int.immigrant <- fit_model(form.immigrant, dat = immigrant, model = "amce", design = immigrant.design)
est.all.immigrant <- ols_int.immigrant[[2]] %>% 
  select(-c(statistic, p.value)) %>% 
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error
  ) %>% 
  rename_with(~ paste0(.x, ".OLS_interact"), .cols = -c(attribute, term)) %>% 
  left_join(est.all.immigrant, ., by = join_by(attribute, term))

# OLS with clustering and interaction
ols_int_clus.immigrant <- fit_model(form.immigrant, dat = immigrant, model = "amce", design = immigrant.design, cluster = "subject")
est.all.immigrant <- ols_int_clus.immigrant[[2]] %>% 
  select(-c(statistic, p.value)) %>% 
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error
  ) %>% 
  rename_with(~ paste0(.x, ".OLS_interact_clustered"), .cols = -c(attribute, term)) %>% 
  left_join(est.all.immigrant, ., by = join_by(attribute, term))
```

再把 OLS 計算出來的 AMCE 另外儲存。

```{r AME-ols}
AME.immigrant <- est.all.immigrant
```



<!-- Result of OLS. -->

```{r ols-res, include=F}
ols.immigrant[[2]] %>%
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error
  ) %>% 
  mutate_if(is.numeric, ~ round(.x, 3)) %>% 
  kable("html", caption="Result of Linear Regression Model for immigrant Experiment") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
  scroll_box(height = "500pt")
```

比較加入或未加入交互作用的 OLS 模型，並且同時檢驗控制或不控制 cluster-robust standard error 後的 CI：

```{r ols-plot, warning=F, message=F}
ols.plot.immigrant <- est.all.immigrant %>% 
  filter(!grepl(":", term)) %>%  # discard interaction terms
  select(c(attribute, term, contains("OLS"))) %>% 
  select(c(attribute, term, starts_with("estimate."), starts_with("CI."))) %>% 
  pivot_longer(
    cols = -c(attribute, term),
    names_to = c(".value", "model"),
    names_pattern = "(.*)\\.(OLS.*)$"
  ) %>%  
  mutate(
    model = fct_recode(
      model, 
      "OLS w/o clustering" = "OLS", 
      "OLS w/ subject clustered" = "OLS_clustered",
      "OLS w/ interaction\n and w/o clustering" = "OLS_interact",
      "OLS w/ interaction\n and w/ subject clustered" = "OLS_interact_clustered"
    )
  ) %>%
  mutate(model = fct_rev(model)) %>%
  plot_result(., n.mod = 2) +
  labs(title = "Immigrant Experiment Fitted by OLS", x = "AMCE", color="Interaction and Clustering")
  
# ggsave("./1-figure/res-immigrant-OLS.png", ols.plot.immigrant)

ols.plot.immigrant
```

可以發現加入交互作用後，那些有相關的 attributes 的確會有影響。

```{r clean-ols, include=F}
# remove OLS objects
rm(ols_clus.immigrant, ols_int.immigrant, ols_int_clus.immigrant, ols.immigrant, ols.plot.immigrant)
```


### Logistic Regression

```{r logit}
# w/o interaction
logit.immigrant <- fit_model(form.immigrant, dat = immigrant, model = "logit")
logit_clus.immigrant <- fit_model(form.immigrant, dat = immigrant, model = "logit", cluster = "subject")

est.all.immigrant <- logit.immigrant[[2]] %>% 
  select(-c(statistic, p.value)) %>% 
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error
  ) %>% 
  rename_with(~ paste0(.x, ".logit"), .cols = -c(attribute, term)) %>% 
  left_join(est.all.immigrant, ., by = join_by(attribute, term))
est.all.immigrant <- logit_clus.immigrant[[2]] %>% 
  select(-c(statistic, p.value)) %>% 
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error
  ) %>% 
  rename_with(~ paste0(.x, ".logit_clustered"), .cols = -c(attribute, term)) %>% 
  left_join(est.all.immigrant, ., by = join_by(attribute, term))

# w/ interaction
logit_int.immigrant <- fit_model(form_int.immigrant, dat = immigrant, model = "logit")
logit_int_clus.immigrant <- fit_model(form_int.immigrant, dat = immigrant, model = "logit", cluster = "subject")

est.all.immigrant <- logit_int.immigrant[[2]] %>% 
  select(-c(statistic, p.value)) %>% 
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error
  ) %>% 
  rename_with(~ paste0(.x, ".logit_interact"), .cols = -c(attribute, term)) %>% 
  right_join(est.all.immigrant, ., by = join_by(attribute, term))
est.all.immigrant <- logit_int_clus.immigrant[[2]] %>% 
  select(-c(statistic, p.value)) %>% 
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error
  ) %>% 
  rename_with(~ paste0(.x, ".logit_interact_clustered"), .cols = -c(attribute, term)) %>% 
  left_join(est.all.immigrant, ., by = join_by(attribute, term))
```

<!-- Result of logistic regression. -->

```{r logit-res, include=F}
logit.immigrant[[2]] %>%
  mutate_if(is.numeric, ~ round(.x, 3)) %>% 
  kable("html", caption="Result of Logit Regression Model for immigrant Experiment") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
  scroll_box(height = "500pt")
```


Plot the result of logistic regression.

```{r logit-plot, warning=F, message=F}
logit.plot.immigrant <- est.all.immigrant %>% 
  select(c(
    attribute, term, (
      contains("logit") & (starts_with("estimate.") | starts_with("CI."))
    )
  )) %>% 
  filter(!grepl(":", term)) %>% 
  mutate(term = factor(term, levels = rev(term))) %>%
  pivot_longer(
    cols = -c(attribute, term),
    names_to = c(".value", "model"),
    names_pattern = "(.*)\\.(logit.*)$"
  ) %>% 
  mutate(model = fct_rev(model)) %>%
  # mutate(
  #   model = fct_recode(
  #     model,
  #     "Logistic regression\n without clustering" = "logit",
  #     "Logistic regression\n with subject clustered" = "logit_clustered",
  #     "Logistic regression\n without clustering" = "logit_interact",
  #     "Logistic regression\n with subject clustered" = "logit_clustered",
  #   )
  # ) %>%
  plot_result(., n.mod = 2) +
  labs(title = "Immigrant Experiment Fitted by Logistic Regression", x = "Estimated Effects", color="Clustering")

# ggsave("./1-figure/res-immigrant-logit.png", logit.plot.immigrant)
logit.plot.immigrant
```


#### AME of Logistic Regression

分別計算有無交互作用模型的 average marginal effect (AME)。

```{r AME-logit, echo=F}
# AME of logistic without interaction and clustering
if (file.exists("./2-res/AME-immigrant-logit.rds")) {
  cat("Found the file of AME of Logistic without interaction and clustering!\n
      Reading file...\n\n")
  AME.logit <- read_rds("./2-res/AME-immigrant-logit.rds")  # read previous result
} else {
  cat("Did not find the file of AME of Logistic without interaction and clustering!\n
      Computing...\n\n")
  # compute AME w/o cluster-robust se
  AME.logit <- avg_comparisons(model = logit.immigrant[[1]], vcov = T)
  write_rds(AME.logit, "./2-res/AME-immigrant-logit.rds")
}

# AME of logistic without interaction and with clustering
if (file.exists("./2-res/AME-immigrant-logit-clustered.rds")) {
  cat("Found the file of AME of Logistic without interaction and with clustering!\n
      Reading file...\n\n")
  AME.logit_clus <- read_rds("./2-res/AME-immigrant-logit-clustered.rds")  # read previous result
} else {
  cat("Did not find the file of AME of Logistic without interaction and with clustering!\n
      Computing...\n\n")
    # compute AME w/ cluster-robust se
  vcov.logit <- vcovCL(logit.immigrant[[1]], cluster = immigrant$subject)
  AME.logit_clus <- avg_comparisons(model = logit.immigrant[[1]], vcov = vcov.logit)
  write_rds(AME.logit_clus, "./2-res/AME-immigrant-logit-clustered.rds")
}


AME.immigrant <- AME.logit %>% 
  as.data.frame() %>%
  select(1:4) %>% 
  rename(attribute = term, term = contrast) %>%
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error,
    term = str_remove(term, " - .*"),
    attribute = str_remove(attribute, "[.]")
  ) %>% 
  rename_with(~ paste0(.x, ".logit"), .cols = -c(attribute, term)) %>% 
  left_join(AME.immigrant, ., by = join_by(attribute, term))
AME.immigrant <- AME.logit_clus %>% 
  as.data.frame() %>%
  select(1:4) %>% 
  rename(attribute = term, term = contrast) %>%
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error,
    term = str_remove(term, " - .*"),
    attribute = str_remove(attribute, "[.]")
  ) %>% 
  rename_with(~ paste0(.x, ".logit_clustered"), .cols = -c(attribute, term)) %>% 
  left_join(AME.immigrant, ., by = join_by(attribute, term))

# show the results
AME.logit %>%
  mutate_if(is.numeric, ~ round(.x, 3)) %>% 
  mutate(term = gsub("*[.]", "", term)) %>% 
  kable("html", caption="AME of Logistic Regression Model for immigrant Experiment") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
  scroll_box(height = "500pt")
AME.logit_clus %>%
  mutate_if(is.numeric, ~ round(.x, 3)) %>% 
  mutate(term = gsub("*[.]", "", term)) %>% 
  kable("html", caption="AME of Logistic Regression Model with Subject Clustered for immigrant Experiment") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
  scroll_box(height = "500pt")
```


接下來是有加入交互作用的 logistic regression model。

```{r AME-logit-interact, echo=F}
AME.logit_int <- compute_immigrant_AME(
  mod = logit_int.immigrant[[1]], 
  file = "./2-res/AME-immigrant-logit-interact.rds", 
  is_cluster = F
)
AME.logit_int_clus <- compute_immigrant_AME(
  mod = logit_int.immigrant[[1]],
  file = "./2-res/AME-immigrant-logit-interact-clustered.rds",
  is_cluster = T
)

# store computed AME
AME.immigrant <- AME.logit_int %>% 
  as.data.frame() %>%
  select(1:4) %>% 
  rename(attribute = term, term = contrast) %>%
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error,
    term = str_remove(term, " - .*"),
    attribute = str_remove(attribute, "[.]")
  ) %>% 
  rename_with(~ paste0(.x, ".logit_interact"), .cols = -c(attribute, term)) %>% 
  left_join(AME.immigrant, ., by = join_by(attribute, term))
AME.immigrant <- AME.logit_int_clus %>% 
  as.data.frame() %>%
  select(1:4) %>% 
  rename(attribute = term, term = contrast) %>%
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error,
    term = str_remove(term, " - .*"),
    attribute = str_remove(attribute, "[.]")
  ) %>% 
  rename_with(~ paste0(.x, ".logit_interact_clustered"), .cols = -c(attribute, term)) %>% 
  left_join(AME.immigrant, ., by = join_by(attribute, term))

# show the results
AME.logit_int %>%
  mutate_if(is.numeric, ~ round(.x, 3)) %>% 
  mutate(term = gsub("*[.]", "", term)) %>% 
  kable("html", caption="AME of Logistic Regression Model for immigrant Experiment") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
  scroll_box(height = "500pt")
AME.logit_int_clus %>%
  mutate_if(is.numeric, ~ round(.x, 3)) %>% 
  mutate(term = gsub("*[.]", "", term)) %>% 
  kable("html", caption="AME of Logistic Regression Model for immigrant Experiment") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
  scroll_box(height = "500pt")
```


<!-- ```{r AME-OLS-interact, cache=T} -->
<!-- AME.OLS_int <- compute_immigrant_AME( -->
<!--   mod = glm(form_int.immigrant, data=immigrant),  -->
<!--   file = "./2-res/AME-immigrant-OLS-interact.rds",  -->
<!--   is_cluster = F -->
<!-- ) -->
<!-- AME.OLS_int_clus <- compute_immigrant_AME( -->
<!--   mod = glm(form_int.immigrant, data=immigrant),  -->
<!--   file = "./2-res/AME-immigrant-OLS-interact-clustered.rds",  -->
<!--   is_cluster = T -->
<!-- ) -->

<!-- # store computed AME -->
<!-- AME.immigrant <- AME.OLS_int %>%  -->
<!--   as.data.frame() %>% -->
<!--   select(1:4) %>%  -->
<!--   rename(attribute = term, term = contrast) %>% -->
<!--   mutate( -->
<!--     CI.lower = estimate - 1.96*std.error, -->
<!--     CI.upper = estimate + 1.96*std.error, -->
<!--     term = str_remove(term, " - .*"), -->
<!--     attribute = str_remove(attribute, "[.]") -->
<!--   ) %>%  -->
<!--   rename_with(~ paste0(.x, ".OLS_glm_interact"), .cols = -c(attribute, term)) %>%  -->
<!--   left_join(AME.immigrant, ., by = join_by(attribute, term)) -->
<!-- AME.immigrant <- AME.OLS_int_clus %>%  -->
<!--   as.data.frame() %>% -->
<!--   select(1:4) %>%  -->
<!--   rename(attribute = term, term = contrast) %>% -->
<!--   mutate( -->
<!--     CI.lower = estimate - 1.96*std.error, -->
<!--     CI.upper = estimate + 1.96*std.error, -->
<!--     term = str_remove(term, " - .*"), -->
<!--     attribute = str_remove(attribute, "[.]") -->
<!--   ) %>%  -->
<!--   rename_with(~ paste0(.x, ".OLS_glm_interact_clustered"), .cols = -c(attribute, term)) %>%  -->
<!--   left_join(AME.immigrant, ., by = join_by(attribute, term)) -->

<!-- # show the results -->
<!-- AME.OLS_int %>% -->
<!--   mutate_if(is.numeric, ~ round(.x, 3)) %>%  -->
<!--   mutate(term = gsub("*[.]", "", term)) %>%  -->
<!--   kable("html", caption="AME of Logistic Regression Model for immigrant Experiment") %>%  -->
<!--   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>% -->
<!--   scroll_box(height = "500pt") -->
<!-- AME.OLS_int_clus %>% -->
<!--   mutate_if(is.numeric, ~ round(.x, 3)) %>%  -->
<!--   mutate(term = gsub("*[.]", "", term)) %>%  -->
<!--   kable("html", caption="AME of Logistic Regression Model for immigrant Experiment") %>%  -->
<!--   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>% -->
<!--   scroll_box(height = "500pt") -->
<!-- ``` -->


```{r clean-logit, include=F}
# remove logit objects
rm(
  logit.immigrant, logit_clus.immigrant, logit_int.immigrant, logit_int_clus.immigrant, 
  AME.logit, AME.logit_clus, AME.logit_int, AME.logit_int_clus,
  logit.plot.immigrant, vcov.logit
)
```


### Conditional Logistic Regression

```{r CL}
# original conditional logistic regression
cl.immigrant <- fit_model(
  update(form.immigrant, . ~ . + strata(trial_id)),
  dat = immigrant, model = "clogit"
)
cl_clus.immigrant <- fit_model(
  update(form.immigrant, . ~ . + strata(trial_id)),
  dat = immigrant, model = "clogit", is_cluster = T
)

est.all.immigrant <- cl.immigrant[[2]] %>% 
  select(-c(statistic, p.value)) %>% 
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error
  ) %>% 
  rename_with(~ paste0(.x, ".CL"), .cols = -c(attribute, term)) %>% 
  left_join(est.all.immigrant, ., by = join_by(attribute, term))
est.all.immigrant <- cl_clus.immigrant[[2]] %>% 
  select(-c(statistic, p.value)) %>% 
  mutate(
    std.error = robust.se,
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error
  ) %>% 
  select(-robust.se) %>% 
  rename_with(~ paste0(.x, ".CL_clustered"), .cols = -c(attribute, term)) %>% 
  left_join(est.all.immigrant, ., by = join_by(attribute, term))
```


```{r CL-int}
cl_int.immigrant <- fit_model(
  update(form_int.immigrant, . ~ . + strata(trial_id)),
  dat = immigrant, model = "clogit"
)
cl_int_clus.immigrant <- fit_model(
  update(form_int.immigrant, . ~ . + strata(trial_id)),
  dat = immigrant, model = "clogit", is_cluster = T
)

est.all.immigrant <- cl_int.immigrant[[2]] %>% 
  select(-c(statistic, p.value)) %>% 
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error
  ) %>% 
  rename_with(~ paste0(.x, ".CL_interact"), .cols = -c(attribute, term)) %>% 
  left_join(est.all.immigrant, ., by = join_by(attribute, term))
est.all.immigrant <- cl_int_clus.immigrant[[2]] %>% 
  select(-c(statistic, p.value, std.error)) %>% 
  rename(std.error = robust.se) %>% 
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error
  ) %>% 
  rename_with(~ paste0(.x, ".CL_interact_clustered"), .cols = -c(attribute, term)) %>% 
  left_join(est.all.immigrant, ., by = join_by(attribute, term))
```



``` {r CL-res, include=F}
cl.immigrant[[2]] %>% 
  mutate_if(is.numeric, ~ round(.x, 3)) %>% 
  kable("html", caption="Result of Conditional Logistic Regression") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>%
  scroll_box(height = "500pt")
```


```{r CL-plot, message=F}
cl.plot.immigrant <- est.all.immigrant %>% 
  select(c(
    attribute, term, (
      contains("CL", ignore.case = F) & (starts_with("estimate.") | starts_with("CI."))
    )
  )) %>% 
  filter(!grepl(":", term)) %>%  # discard interaction terms
  mutate(term = factor(term, levels = rev(term))) %>% 
  pivot_longer(
    cols = -c(attribute, term),
    names_to = c(".value", "model"),
    names_pattern = "(.*)\\.(CL.*)$"
  ) %>%  
  mutate(
    model = fct_recode(
      model, 
      "Conditional logistic\n w/o clustering" = "CL",
      "Conditional logistic\n w/ subject clustered" = "CL_clustered",
      "Conditional logistic\n w/ interaction \n w/o subject clustered" = "CL_interact",
      "Conditional logistic\n w/ interaction \n w/ subject clustered" = "CL_interact_clustered",
    )
  ) %>%
  mutate(model = fct_rev(model)) %>%
  plot_result(., n.mod = 4) +
  labs(
    title = "immigrant Experiment Fitted by Condtional Logistic Regression",
    x = "Estimated Effects", color = "Interaction and Clustering"
  )

ggsave("./1-figure/res-immigrant-CL.png", cl.plot.immigrant)
cl.plot.immigrant
```


#### AME of Conditional Logistic Regression

```{r AME-CL, echo=F}
# model & covariance matrice objects
cl.mod <- clogit(update(form.immigrant, . ~ . + strata(trial_id)), data = immigrant)
cl_clus.mod <- clogit(update(form.immigrant, . ~ . + strata(trial_id)), data = immigrant, method = "efron", cluster = subject)
cl.vcov <- vcov(cl.mod)
cl.vcovCL <- vcovCL(cl_clus.mod, immigrant$subject)

# compute AME w/ both std.error & cluster-robust std.error
AME.cl <- compute_immigrant_AME.cl(cl.mod, immigrant, "./2-res/AME-immigrant-CL.rds", cl.vcov, cl.vcovCL)

# store computed AME
AME.immigrant <- AME.cl %>% 
  as.data.frame() %>%
  select(1:4) %>% 
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error,
    attribute = str_remove(attribute, "[.]")
  ) %>% 
  rename_with(~ paste0(.x, ".CL"), .cols = -c(attribute, term)) %>% 
  left_join(AME.immigrant, ., by = join_by(attribute, term))
AME.immigrant <- AME.cl %>% 
  as.data.frame() %>%
  select(c(1:3, 5)) %>% 
  rename(std.error = robust.se) %>%
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error,
    attribute = str_remove(attribute, "[.]")
  ) %>% 
  rename_with(~ paste0(.x, ".CL_clustered"), .cols = -c(attribute, term)) %>% 
  left_join(AME.immigrant, ., by = join_by(attribute, term))
```


接下來是有加入交互作用的 conditional logistic regression model。

```{r AME-CL-interact, echo=F}
# get model and covariance matrice object
cl_int.mod <- clogit(update(form_int.immigrant, . ~ . + strata(trial_id)), data = immigrant)
cl_int_clus.mod <- clogit(update(form_int.immigrant, . ~ . + strata(trial_id)), data = immigrant, method = "efron", cluster = subject)
cl_int.vcov <- vcov(cl_int.mod)
cl_int.vcovCL <- vcovCL(cl_int_clus.mod, immigrant$subject)

# compute AME with both std.error and cluster-robust std.error
AME.cl_int <- compute_immigrant_AME.cl(cl_int.mod, immigrant, "./2-res/AME-immigrant-CL-interact.rds", cl_int.vcov, cl_int.vcovCL)

# store computed AME
AME.immigrant <- AME.cl_int %>% 
  as.data.frame() %>%
  select(1:4) %>% 
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error,
    attribute = str_remove(attribute, "[.]")
  ) %>% 
  rename_with(~ paste0(.x, ".CL_interact"), .cols = -c(attribute, term)) %>% 
  left_join(AME.immigrant, ., by = join_by(attribute, term))
AME.immigrant <- AME.cl_int %>% 
  as.data.frame() %>%
  select(c(1:3, 5)) %>% 
  rename(std.error = robust.se) %>%
  mutate(
    CI.lower = estimate - 1.96*std.error,
    CI.upper = estimate + 1.96*std.error,
    attribute = str_remove(attribute, "[.]")
  ) %>% 
  rename_with(~ paste0(.x, ".CL_interact_clustered"), .cols = -c(attribute, term)) %>% 
  left_join(AME.immigrant, ., by = join_by(attribute, term))
```


```{r AME-plot-CL}
AME.plot.cl <- AME.immigrant %>% 
  select(c(attribute, term, contains("CL", ignore.case=F))) %>%
  mutate(term = factor(term, levels = rev(term))) %>% 
  pivot_longer(
    cols = -c(attribute, term),
    names_to = c(".value", "model"),
    names_pattern = "(.*)\\.(.*)$"
  ) %>%
  mutate(
    model = fct_recode(
      model,
      "Conditional logistic\n w/o clustering" = "CL",
      "Conditional logistic\n w/ subject clustered" = "CL_clustered",
      "Conditional logistic\n w/ interaction\n w/o clustering" = "CL_interact",
      "Conditional logistic\n w/ interaction\n and subject clustered" = "CL_interact_clustered",
    )
  ) %>%
  mutate(model = fct_rev(model)) %>%
  plot_result(., n.mod = 4) +
  labs(
    title = "AME of Immigrant Experiment Fitted by Conditional Logistic Regression", 
    x = "Estimated Effects", color = "Model"
  )

# ggsave("./1-figure/AME-immigrant-CL.png", AME.plot.cl)
AME.plot.cl
```


```{r clean-CL, warning=F, include=F}
# remove CL objects
rm(
  cl.immigrant, cl_clus.immigrant, cl_int.immigrant, cl_int_clus.immigrant, 
  cl.mod, cl_clus.mod, cl_int.mod, cl_int_clus.mod,
  cl.vcov, cl.vcovCL, cl_int.vcov, cl_int.vcovCL,
  AME.cl, AME.cl_int,
  cl.plot.immigrant, AME.plot.cl
)
```


## Result Comparison

### Logistic and Conditional Logistic Regression w/ and w/o Clustering

```{r plot-all, fig.height=12, fig.width=7}
est.all.immigrant %>% 
  select(-contains("OLS")) %>%
  filter(!grepl(":", term)) %>% 
  mutate(term = factor(term, levels = rev(term))) %>% 
  pivot_longer(
    cols = -c(attribute,term),
    names_to = c(".value", "model"),
    names_pattern = "(.*)\\.(.*)$"
  ) %>%  
  # mutate(
  #   model = fct_recode(
  #     model,
  #     "Logistic regression\nwithout clustering" = "logit",
  #     "Logistic regression\nwith subject clustered" = "logit_clustered",
  #     "Conditional logistic\nwithout clustering" = "CL",
  #     "Conditional logistic\nwith subject clustered" = "CL_clustered"
  #   )
  # ) %>% 
  mutate(model = fct_rev(model)) %>%
  plot_result(., n.mod=4) +
  labs(
    x = "Effect estimate", col="Model",
    title = "Estimated Effects in Immigrant Experiment: \nComparison Between Logistic and Conditional Logistic"
  ) +
  theme(
      # legend.position = "bottom",
      plot.title.position = "plot"
    )
# ggsave("./1-figure/res-immigrant-BL+CL.png")
```

<!-- Robust se 的膨脹程度通常可以由 Moulton factor 做近似: -->

<!-- $$ -->
<!-- \frac{SE_{robust}}{SE_{standard}} \approx \sqrt{1+(n-1)\rho_x \rho_\varepsilon} -->
<!-- $$ -->

<!-- ，其中 $n$ 是 cluster size，$\pho_x$ 是 attributes 在 cluster 內部的相關性，而 $\rho_\varepsilon$ 則是殘差的相關性。 -->

<!-- 由於 conjoint experiment 的 randomization 設計，使得每位受試者觀察到的 attributes 之間不存在相關性， -->
<!-- 因此 Moulton factor 會約為 1，說明了原本的 standard error 和 cluster-robust standard error 之間為何差異不大。 -->


### AME Comparisons

```{r AME-plot-all, fig.height=12, fig.width=7}
AME.plot.all <- AME.immigrant %>% 
  select(c(attribute, term, contains("interact"))) %>%
  mutate(term = factor(term, levels = rev(term))) %>% 
  pivot_longer(
    cols = -c(attribute, term),
    names_to = c(".value", "model"),
    names_pattern = "(.*)\\.(.*)$"
  ) %>%
  # mutate(
  #   model = fct_recode(
  #     model,
  #     "Conditional logistic\n w/o clustering" = "CL",
  #     "Conditional logistic\n w/ subject clustered" = "CL_clustered",
  #     "Conditional logistic\n w/ interaction\n w/o clustering" = "CL_interact",
  #     "Conditional logistic\n w/ interaction\n and subject clustered" = "CL_interact_clustered",
  #   )
  # ) %>%
  mutate(model = fct_rev(model)) %>%
  plot_result(., n.mod = 4) +
  labs(
    title = "AME of Immigrant Experiment Fitted by 3 Models", 
    x = "Estimated Effects", color = "Model"
  )

# ggsave("./1-figure/AME-immigrant-all.png", AME.plot.all)
AME.plot.all
```



