---
title: "Compare the Estimations of Logit and Conditional Logit"
author: "Wei-Lun, Lin"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)  # data processing
library(broom)
library(kableExtra)  # print table
library(survival)
library(cjoint)
# library(marginaleffects)

library(numDeriv)
library(sandwich)
library(lmtest)
library(multiwayvcov)

source("./0-code/functions.R")
```

-   背景: 由於 logit model 和 conditional logit model 很像，因此想要知道在何種情形之下，這兩種模型是等同的。 <!-- -   結果: 比較 logistic 和 conditional logistic，發現 --> <!-- ... OLS 得到的 AMCE 和 logistic regression 得到的估計非常接近，而 logistic 和 conditional logistic regression 的估計結果也相差不大。此外，分別比較 logistic 和 conditional logistic regression 可以發現，控制 cluster-robust standard error 前後得到的 standard error 也很類似，此結果和實驗的隨機生成 profiles 機制息息相關。 --> <!-- -   未來工作:... 再加入無母數的方法，先對受試者的潛在分群特徵做分群，再用模型分別 fit 資料，並比較事先和事後控制 cluster 所得到的表現。 -->

## Models

Under conjoint experiment design, subject $i$ is provided with $J$ alternatives, which composed of $L$ attributes, and he/she has to choose between the alternatives.

### Conditional Logit Model

When there are only 2 alternatives in a trial, the probability of subject $i$ choosing alternative $1$ at trial $k$ is

$$
\Pr(Y_{i1k}=1)=
  \frac
    {\exp(\boldsymbol{X_{i1k}\alpha})}
    {\exp(\boldsymbol{X_{i1k}\alpha})+\exp(\boldsymbol{X_{i2k}\alpha)}}
$$

, where $Y_{ijk}\in \{0, 1\}$ denotes the chosen or not and the vector $\boldsymbol{X_{ijk}}\in \mathbb{R}^L$ denotes the values of the corresponding predictors.

### Logit Model

The probability of subject $i$ choosing alternative $j$ at trial $k$ is

$$
\Pr(Y_{ijk}=1)=\frac{1}{1+\exp(-(\beta_0 + \boldsymbol{X_{ijk}\beta}))}
$$

, where $\beta_0$ denotes the intercept and $\boldsymbol{\beta}\in\mathbb{R}^L$ denote the slopes.

## Simulation of Conjoint Experiment with 1 Attribute

### Data Generating Process

Consider a conjoint experiment that there is only 1 attribute: gender: male `0` or female `1`.

In the simulation, we have `5000` subjects, and they all choose between the two alternatives: male and female.

$$
p_i \sim Beta(\alpha=1, \beta=2) \\
Y_i \sim B(n=1, p_i)
$$

```{r dat L1}
n <- 5000
set.seed(322)

dat.L1 <- data.frame(
  subject = 1:n,
  trial = 1:n,
  p1 = rbeta(n, 1, 2),
  gender = 0,
  alternative = 1
) %>% 
  rowwise() %>% 
  mutate(select = rbinom(1, 1, p1)) %>% 
  ungroup()
dat.L1 <- dat.L1 %>% 
  mutate(
    select = 1 - select,
    alternative = 2,
    gender = 1
  ) %>% 
  bind_rows(dat.L1, .) %>% 
  arrange(subject, alternative) %>% 
  mutate(id = row_number(), .before = subject)
dat.L1 %>% 
  head() %>% 
  kable("html", caption="Data with 1 Attribute") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```

### Data Fitting

```{r logit L1}
logit.L1 <- glm(select ~ gender, family = binomial(link = "logit"), data = dat.L1)
summary(logit.L1)
```

```{r clogit L1}
cl.L1 <- clogit(select ~ gender + strata(trial), data = dat.L1)
summary(cl.L1)
```

### Result

The coefficients are

|   | Intercept | Gender |
|------------------|------------------------------------|------------------|
| Logit | $\beta_0=`r coef(logit.L1)[[1]] %>% round(3)`$ | $\beta_1=`r coef(logit.L1)[[2]] %>% round(3)`$ |
| Conditional logit | NA | $\alpha=`r coef(cl.L1)[[1]] %>% round(3)`$ |

It turns out that the slope of conditional logit model equals to the negative intercept of logit model ($-\alpha=\beta_0$), and also equals to the half of the slope of logit model ($\alpha=\beta_1/2$).

When there is only 1 attribute, the predictors are $X_{i1k}=0$ and $X_{i2k}=1$, and there is only 1 trial. In conditional logit model, the the probability of subject $i$ choosing alternative $1$ be

$$
\begin{align*}
  P_{i1, CL}
    &= \frac{\exp(0)}{\exp(0)+\exp(\alpha)} \\
    &= \frac{1}{1+\exp(\alpha)},
\end{align*}
$$

and that of choosing alternative $2$ be

$$
P_{i2, CL} = \frac{\exp(\alpha)}{1+\exp(\alpha)} = 1 - P_{i1k,CL}.
$$

Let alternative $1$ be chosen $n_1$ times and alternative $2$ be chosen $n_2$ times in the whole dataset. The log likelihood function is

$$
\begin{align*}
  l(\alpha)
    &= \sum^n_{i=1}[Y_{i1}\cdot\ln(P_{i1,CL})+Y_{i2}\cdot \ln(P_{i2,CL})] \\
    &= n_1\ln(P_{i1,CL})+n_2\ln(1-P_{i1,CL}).
\end{align*}
$$

To perform maximum likelihood estimation, we set the first derivative of the log likelihood function to 0.

$$
\begin{align*}
  \frac{\partial l(\alpha)}{\partial\alpha}
    &= \frac{\partial l(\alpha)}{\partial P_{i1,CL}} \frac{\partial P_{i1,CL}}{\partial\alpha} = 0,
\end{align*}
$$

then $\hat P_{i1,CL}=\frac{n_1}{n_1+n_2}$ and $\hat\alpha=\ln({n_2\over n_1})$.

In logit model, the probability of subject $i$ choosing alternative $1$ is

$$
\begin{align*}
  P_{i1, L}
    &= \frac{1}{1+\exp(-\beta_0-0\cdot\beta_1)} \\
    &= \frac{1}{1+\exp(-\beta_0)},
\end{align*}
$$

and that of choosing alternative $2$ is

$$
\begin{align*}
  P_{i2, L}
    &= \frac{1}{1+\exp(-(\beta_0 + \beta_1))}.
\end{align*}
$$

The log likelihood function is

$$
\begin{align*}
  l(\beta_0, \beta_1)
    &= \sum^n_{i=1}\sum^2_{j=1}[Y_{ij}\ln(P_{ij, L})+(1-Y_{ij})\ln(1-P_{ij, L})] \\
    &= n_1\ln(P_{i1, L}) + n_2\ln(1-P_{i1, L}) + n_2\ln(P_{i2, L}) + n_1\ln(1-P_{i2, L}).
\end{align*}
$$

Set the two partial derivatives to zero, and we got the maximum likelihood estimator.

$$
\begin{cases}
  \hat\beta_0 = -\ln({n_2\over n_1}) = -\hat\alpha, \\
  \hat\beta_1 = 2\ln({n_2\over n_1}) = 2\hat\alpha.
\end{cases}
$$

## Simulation of Conjoint Experiment with 2 Attributes

In this simulation, there are 2 binary variables: + gender: male `0` or female `1` + military: served `1` or not `0`

Thus, there are *4* potential profiles. To compare all pairs of combinations, there should be ${4\choose2}=6$ trials per subject.

```{r design L2}
design.2 <- expand.grid(gender = 0:1, military = 0:1)
ind <- combn(nrow(design.2), 2)
trials.2 <- bind_rows(
  design.2[ind[, 1], ]
)

for (i in 2:ncol(ind)) {
  trials.2 <- bind_rows(trials.2, design.2[ind[, i],])
}

n_sub <- 2000
dat.L2 <- trials.2[rep(seq_len(nrow(trials.2)), n_sub), ]
row.names(dat.L2) <- NULL
dat.L2 <- dat.L2 %>% 
  mutate(
    row.names = NULL,
    id = row_number(),
    subject = rep(1:n_sub, each=12),
    trial = rep(1:(n_sub * 6), each=2),
    .before = gender
  )
```

```{r dat L2}
b0 <- -1.2
b1 <- 1.1
b2 <- 0.8

dat.L2 <- dat.L2 %>% 
  group_by(trial) %>% 
  mutate(
    xb = b0 + b1 * gender + b2 * military,
    p1 = 1 / (1 + exp(-(xb[1]-xb[2]))),
    select = rbinom(1, 1, p1),
    select = ifelse(row_number() == 1, select, 1 - select)
  ) %>% 
  ungroup() %>% 
  select(-xb)
dat.L2 %>% 
  head(24) %>% 
  kable("html", caption="Data with 2 Attributes (First 24 Rows)") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) %>% 
  scroll_box(width = "100%")
```

### Data Fitting

```{r logit L2}
logit.L2 <- glm(select ~ gender + military, family = binomial(), data = dat.L2)
summary(logit.L2)
```

```{r clogit L2}
cl.L2 <- clogit(select ~ gender + military + strata(trial), data = dat.L2)
summary(cl.L2)
```

The coefficients are

|   | Intercept | Gender | Military |
|---------------|---------------------------|---------------|---------------|
| Logit | $\beta_0=`r coef(logit.L2)[[1]] %>% round(2)`$ | $\beta_1=`r coef(logit.L2)[[2]] %>% round(2)`$ | $\beta_2=`r coef(logit.L2)[[3]] %>% round(2)`$ |
| Conditional logit | NA | $\alpha_1=`r coef(cl.L2)[[1]] %>% round(2)`$ | $\alpha_2=`r coef(cl.L2)[[2]] %>% round(2)`$ |
